Modeling the Latent Space of Symbolic String Quartet Music

### 2021-06-08 papers/final/alex/paper.tex
metric analysis in paper

Data augmentation was also tested--Oore et al (2018)
\cite{oore_this_2018} suggests augmentation via pitch shifting each
training example up or down by up to six semitones in order to create
additional training examples and reduce overfitting, so we followed
that recommendation and used it to double the size of the training
dataset. As a result, we saw much lower validation losses (compare
Figure \ref{loss} and Figure \ref{loss_augment} and a small but
noticeable improvement in the fidelity of the song reconstructions,
suggesting that this method was effective in reducing overfitting.

@@

The reconstructed music produced by the model sounds reminiscent of
the original inputs, preserving the key signature and general rhythmic
and melodic structure, albeit with noticeable differences at the
individual note level.

@@

The interpolations are between the following 18 pairs of pieces

\begin{itemize}
\tightlist
\item \texttt{Haydn/2104\_op64n5\_1.mid} , \texttt{Ravel/2178\_gr\_rqtf2.mid}
\item \texttt{Haydn/2105\_op64n5\_2.mid} ,   \texttt{Ravel/2179\_gr\_rqtf3.mid}
\item \texttt{Haydn/2106\_op64n5\_3.mid} ,   \texttt{Ravel/2177\_gr\_rqtf1.mid}
\item \texttt{Beethoven/2497\_qt11\_4.mid} , \texttt{Mozart/1822\_kv\_421\_1.mid}
\item \texttt{Beethoven/2433\_qt16\_3.mid} , \texttt{Mozart/1859\_kv\_464\_2.mid}
\item \texttt{Beethoven/2368\_qt12\_4.mid} , \texttt{Mozart/1807\_kv\_387\_3.mid}
\item \texttt{Beethoven/2314\_qt15\_2.mid} , \texttt{Mozart/1791\_kv\_465\_4.mid}
\item \texttt{Beethoven/2480\_qt05\_1.mid} , \texttt{Mozart/1792\_kv\_465\_1.mid}
\item \texttt{Beethoven/2481\_qt05\_2.mid} , \texttt{Mozart/1835\_kv\_590\_3.mid}
\item \texttt{Beethoven/2379\_qt08\_4.mid} , \texttt{Mozart/1805\_kv\_387\_1.mid}
\item \texttt{Beethoven/2365\_qt12\_1.mid} , \texttt{Mozart/1793\_kv\_465\_2.mid}
\item \texttt{Beethoven/2562\_qt02\_4.mid} , \texttt{Mozart/1790\_kv\_465\_3.mid}
\item \texttt{Beethoven/2494\_qt11\_1.mid} , \texttt{Mozart/1789\_kv\_465\_2.mid}
\item \texttt{Beethoven/2403\_qt01\_4.mid} , \texttt{Mozart/1788\_kv\_465\_1.mid}
\item \texttt{Beethoven/2376\_qt08\_1.mid} , \texttt{Dvorak/1916\_dvq10m1.mid}
\item \texttt{Beethoven/2384\_qt13\_4.mid} , \texttt{Bach/2242\_vs1\_2.mid}
\item \texttt{Beethoven/2560\_qt02\_2.mid} , \texttt{Beethoven/2621\_qt07\_1.mid}
\item \texttt{Beethoven/2377\_qt08\_2.mid} , \texttt{Beethoven/2381\_qt13\_1.mid}
\end{itemize}

\subsection{Model Evaluation}

Evaluation of generative models is challenging because there is no
equivalent of an accuracy metric like what is used in supervised
learning. For autoencoder models we can measure how accurately the
model can reconstruct its own inputs, but this does not tell us the
quality of the interpolated examples. Generative models are typically
evaluated using a combination of qualitative metrics whereby human
judges rate the quality of the generated examples (essentially a
Turing test), and quantitative metrics that assess the differences in
the parametric distributions of generated and real examples. Yang and
Lerch (2020) proposes a set of metrics informed by music theory, for
probabilistically evaluating how similar the generations are to known
sample distributions of real music \cite{yang_evaluation_2020}. These
metrics include counts, ranges, histograms and transition matrices of
pitches and note lengths, then the Kullback-Leibler divergence and
overlapping area of the probability density functions are used to
compare against known reference distributions per musical genre
\cite{yang_evaluation_2020}. Due to the cost and time requirements
associated with designing a human subjects experiment, we
utilize this quantitative approach to the quality assessment of
generated samples.

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.6}{\input{figures/source_pitch_count.pgf}}
    \end{center}
    \caption{Distinct pitch counts per sample for source samples.}
    \label{source_pitches}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.6}{\input{figures/generated_pitch_count.pgf}}
    \end{center}
    \caption{Distinct pitch counts per sample for generated samples.}
    \label{generated_pitches}
\end{figure}

>>>>

Data augmentation was also tested--Oore et al (2018)

\cite{oore_this_2018} suggests augmentation via pitch shifting each
training example up or down by up to six semitones in order to create
additional training examples and reduce overfitting, so we followed
that recommendation and used it to double the size of the training
dataset. As a result, we saw much lower validation losses (compare
Figure \ref{loss} and Figure \ref{loss_augment} and a small but
noticeable improvement in the fidelity of the song reconstructions,
suggesting that this method was effective in reducing overfitting.

@@

The reconstructed music produced by the model sounds reminiscent of
the original inputs, preserving the key signature and general rhythmic
and melodic structure, albeit with small but noticeable differences at
the individual note level, such as more silence and less variation in
pitch.

@@

The interpolations are between the following 18 pairs of pieces:

\begin{itemize}
\tightlist
\item \texttt{Haydn/2104\_op64n5\_1.mid} \& \texttt{Ravel/2178\_gr\_rqtf2.mid}
\item \texttt{Haydn/2105\_op64n5\_2.mid} \&   \texttt{Ravel/2179\_gr\_rqtf3.mid}
\item \texttt{Haydn/2106\_op64n5\_3.mid} \&   \texttt{Ravel/2177\_gr\_rqtf1.mid}
\item \texttt{Beethoven/2497\_qt11\_4.mid} \& \texttt{Mozart/1822\_kv\_421\_1.mid}
\item \texttt{Beethoven/2433\_qt16\_3.mid} \& \texttt{Mozart/1859\_kv\_464\_2.mid}
\item \texttt{Beethoven/2368\_qt12\_4.mid} \& \texttt{Mozart/1807\_kv\_387\_3.mid}
\item \texttt{Beethoven/2314\_qt15\_2.mid} \& \texttt{Mozart/1791\_kv\_465\_4.mid}
\item \texttt{Beethoven/2480\_qt05\_1.mid} \& \texttt{Mozart/1792\_kv\_465\_1.mid}
\item \texttt{Beethoven/2481\_qt05\_2.mid} \& \texttt{Mozart/1835\_kv\_590\_3.mid}
\item \texttt{Beethoven/2379\_qt08\_4.mid} \& \texttt{Mozart/1805\_kv\_387\_1.mid}
\item \texttt{Beethoven/2365\_qt12\_1.mid} \& \texttt{Mozart/1793\_kv\_465\_2.mid}
\item \texttt{Beethoven/2562\_qt02\_4.mid} \& \texttt{Mozart/1790\_kv\_465\_3.mid}
\item \texttt{Beethoven/2494\_qt11\_1.mid} \& \texttt{Mozart/1789\_kv\_465\_2.mid}
\item \texttt{Beethoven/2403\_qt01\_4.mid} \& \texttt{Mozart/1788\_kv\_465\_1.mid}
\item \texttt{Beethoven/2376\_qt08\_1.mid} \& \texttt{Dvorak/1916\_dvq10m1.mid}
\item \texttt{Beethoven/2384\_qt13\_4.mid} \& \texttt{Bach/2242\_vs1\_2.mid}
\item \texttt{Beethoven/2560\_qt02\_2.mid} \& \texttt{Beethoven/2621\_qt07\_1.mid}
\item \texttt{Beethoven/2377\_qt08\_2.mid} \& \texttt{Beethoven/2381\_qt13\_1.mid}
\end{itemize}

\subsection{Generated Sample Evaluation}

Evaluation of generative models is challenging because there is no
direct equivalent of an accuracy or error metric like what is used in
supervised learning. For autoencoder models we can measure how
accurately the model can reconstruct its own inputs, but this does not
tell us the quality of the interpolated examples. Generative models
are typically evaluated using a combination of qualitative metrics
whereby human judges rate the quality of the generated examples and
guess whether they were generated by a machine (essentially a Turing
test), and quantitative metrics that assess the differences in the
parametric distributions of generated and real examples. Yang and
Lerch (2020) propose a set of metrics informed by music theory, for
probabilistically evaluating how similar the generations are to known
sample distributions of real music \cite{yang_evaluation_2020}. These
metrics include counts, ranges, histograms and transition matrices of
pitches and note lengths, then the Kullback-Leibler divergence and
overlapping area of the probability density or mass functions are
used to compare against known reference distributions per musical
genre \cite{yang_evaluation_2020}. Due to the cost and time
requirements associated with designing a human subjects experiment for
qualitative evaluation, we utilize this quantitative approach to the
assessment of generated samples.

Figure \ref{pitch_counts} depicts histograms of source and generated
music samples respectively (1 sample = 1 measure) by the number of
distinct pitch counts they contain. In both samples the mode is 9
distinct pitches, but the source arrangements have a median of 11
while the generations have a median of 8. This supports our own
qualitative observation that the generated samples have slightly less
pitch diversity than the genuine compositions.

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.6}{\input{figures/pitch_count_histogram.pgf}}
    \end{center}
    \caption{Distinct pitch counts per sample for generated and source samples.}
    \label{pitch_counts}
\end{figure}

Figure \ref{pch} depicts the source and generated samples' proportion
of notes by each of the 12 notes in the chromatic scale, regardless of
octave, where 0 = C, 1 = C$\sharp$/D$\flat$, 2 = D, and so on. These
distributions are very similar, with the generated samples having more
Cs and fewer Ds than the source samples.
